---
title: "RWorksheet_camasa#2"
author: "April Marie"
date: "2024-09-25"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1.Create a vector using : operator
# a. Sequence from -5 to 5. Write the R code and its output.
```{r}
operator <- c(-5:5)
print(operator)
```
# The output is sequenced from -5 to 5.

# b. x <- 1:7. What will be the value of x?
```{r}
x <- seq(1:7)
print(x)
```
# The values of x are 1 2 3 4 5 6 7.

# 2. Create a vector using seq() function
# a. seq(1, 3, by=0.2) # specify step size
```{r}
seqC <- seq(1, 3, by=0.2)
print(seqC)
```
# The output is a numeric vector that starts at 1 and ends at 3, with increments (or step size) of 0.2.

# 3. A factory has a census of its workers. There are 50 workers in total. The following
list shows their ages: 34, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27, 22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 43, 53, 41, 51, 35, 24,33, 41, 53, 40, 18, 44, 38, 41, 48, 27, 39, 19, 30, 61, 54, 58, 26, 18.
# a. Access 3rd element, what is the value?
```{r}
ages <- c(34, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27, 22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 43, 53, 41, 51, 35, 24, 33, 41, 53, 40, 18, 44, 38, 41, 48, 27, 39, 19, 30, 61, 54, 58, 26, 18)
third <- ages[3]
print(third)
```
# b. Access 2nd and 4th element, what are the values?
```{r}
second_fourth <- ages[c(2, 4)]
print(second_fourth)
```
# c. Access all but the 4th and 12th element is not included. Write the R script and its output.
```{r}
age <- ages[-c(4, 12)]
print(age)
```
# This output is the original list of ages without the 4th and 12th elements.

# 4. Create a vector x <- c("first"=3, "second"=0, "third"=9). Then named the vector, names(x).
```{r}
names <- c("first"=3, "second"=0, "third"=9)
print(names)
```
# a. Print the results. Then access x[c("first", "third")]. Describe the output.
```{r}
names[c("first", "third")]
```
# The output prints the first and third.

# 5. Create a sequence x from -3:2.
```{r}
seqA <- seq(-3:2)
print(seqA)
```
# a. Modify 2nd element and change it to 0;
#x[2] <- 0 
#x. 
#Describe the output.
```{r}
x[2] <- 0 
x
```

# 6. The following data shows the diesel fuel purchased by Mr. Cruz.
# a. Create a data frame for month, price per liter (php) and purchase-quantity (liter). Write the R scripts and its output.
```{r}
months <- c("January", "February", "March", "April", "May", "June")
price_per_liter <- c(52.50, 57.25, 60.00, 65.00, 74.25, 54.00)
purchase_quantity <- c(25, 30, 40, 50, 10, 45)

data <- data.frame(Month = months, Price_Per_Liter = price_per_liter, Purchase_Quantity = purchase_quantity)
print(data)
```
# b. What is the average fuel expenditure of Mr. Cruz from Jan to June? Note: Use ‘weighted.mean(liter, purchase)‘. Write the R scripts and its output.
```{r}
expenditure <- price_per_liter * purchase_quantity

average_expenditure <- weighted.mean(expenditure, purchase_quantity)

print(average_expenditure)
```

# 7. R has actually lots of built-in datasets. For example, the rivers data “gives the lengths (in miles) of 141 “major” rivers in North America, as compiled by the US Geological Survey”.
# Type “rivers” in your R console.
```{r}
data <- c(length(rivers), sum(rivers), mean(rivers), median(rivers), var(rivers), sd(rivers), min(rivers), max(rivers))
print(data)
```

# 8. The table below gives the 25 most powerful celebrities and their annual pay as ranked by the editions of Forbes magazine and as listed on the Forbes.com website.
# a. Create vectors according to the above table. Write the R scripts and its output.
```{r}
names <- c("Tom Cruise", "Rolling Stones", "Oprah Winfrey", "U2", "Tiger Woods", "Steven Spielberg", "Howard Stern", "50 Cent", "Cast of the Sopranos", "Dan Brown", "Bruce Springsteen", "Donald Trump", "Muhammad Ali", "Paul McCartney", "George Lucas", "Elton John", "David Letterman", "Phil Mickelson", "J.K Rowling", "Bradd Pitt", "Peter Jackson", "Dr. Phil McGraw", "Jay Lenon", "Celine Dion", "Kobe Bryant")
power_ranking <- 1:25
pay <- c(67, 90, 225, 110, 90, 332, 302, 41, 52, 88,
                55, 44, 55, 40, 233, 34, 40, 47, 75, 25,
                39, 45, 32, 40, 31)
celebrity_data <- data.frame(names, power_ranking, pay)
print(celebrity_data)
```
# b. Modify the power ranking and pay of J.K. Rowling. Change power ranking to 15 and pay to 90. Write the R scripts and its output.
```{r}
power_ranking[which(celebrity_name =="J.K. Rowling")] <- 15
pay[which(celebrity_name == "J.K Rowling")] <- 90

