---
title: "RWorksheet_camasa#4b"
author: "April Marie"
date: "2024-10-30"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Using the for loop, create an R script that will display a 5x5 matrix as shown in Figure 1. It must contain vectorA = [1,2,3,4,5] and a 5 x 5 zero matrix.
```{r}
vectorA <- c(1, 2, 3, 4, 5)

zmat <- matrix(0, nrow = 5, ncol = 5)
rmat <- zmat

for (i in 1:5) {
  for (j in 1:5) {
    rmat[i, j] <- abs(vectorA[i] - zmat[i, j])
  }
}
print(rmat)
```

# 2. Print the string "*" using for() function.
```{r}
a <- 5

for (i in 1:a) {
  r_output <- paste(rep("*", i), collapse = " ")
  cat(r_output, "\n")
}
```

# 3. Get an input from the user to print the Fibonacci sequence starting from the 1st input up to 500. Use repeat and break statements.
```{r eval=FALSE}
start <- as.numeric(readline(prompt = "Enter a number: "))

x <- 0
y <- 1

cat("Fibonacci sequence starting from", start, "up to 500:\n")

repeat {
  fib_seq <- x + y
 
  if (fib_seq > 500) {
    break  
  }
 
  if (fib_seq >= start) {
    cat(fib_seq, "\n")
  }
 
  x <- y
  y <- fib_seq
}
```

# 4. Import the dataset as shown in Figure 1 you have created previously.
# a. What is the R script for importing an excel or a csv file? Display the first 6 rows of the dataset?
```{r}
data <- read.csv("shoe_sizes.csv")

head(data)
```

# b. Create a subset for gender(female and male). How many observations are there in Male? How about in Female?
```{r}
fdata <- subset(data, Gender == "F")
mdata <- subset(data, Gender == "M")

num_female <- nrow(fdata)
num_male <- nrow(mdata)

num_female
num_male
```

# c. 
```{r}
genderCount <- table(data$Gender)
barplot(genderCount,
        main = "Number of Males and Females in Household Data",
        col = c("blue", "purple"),
        legend = rownames(genderCount),
        ylab = "Count",
        xlab = "Gender",
        beside = TRUE)
```

# 5. 
```{r}
exp <- c(60, 10, 5, 25)
labels <- c("Food", "Electricity", "Savings", "Miscellaneous")
colors <- c("skyblue", "lightgreen", "beige", "violet")

pie(exp,
    labels = paste(labels, "\n", round((exp / sum(exp)) * 100, 1), "%", sep = ""),
    col = colors,
    main = "Monthly Income Distribution of Dela Cruz Family")

```
# 6.
```{r}
# a.
data(iris)

str(iris)

# b.
meanValues <- colMeans(iris[, 1:4])

print(meanValues)

# c.
speciesCounts <- table(iris$Species)

colors <- c("lightblue", "lightyellow", "lightgreen")

pie(speciesCounts,
    labels = paste(names(speciesCounts), "\n", speciesCounts, sep = ""),
    col = colors,
    main = "Species Distribution in Iris Dataset")

# d.
setosaSubset <- iris[iris$Species == "setosa", ]
versicolorSubset <- iris[iris$Species == "versicolor", ]
virginicaSubset <- iris[iris$Species == "virginica", ]

setosaLastSix <- tail(setosaSubset, 6)
versicolorLastSix <- tail(versicolorSubset, 6)
virginicaLastSix <- tail(virginicaSubset, 6)

print("Last six rows of Setosa:")
print(setosaLastSix)
print("Last six rows of Versicolor:")
print(versicolorLastSix)
print("Last six rows of Virginica:")
print(virginicaLastSix)

# e.
iris$Species <- as.factor(iris$Species)

plot(iris$Sepal.Length, iris$Sepal.Width,
     col = iris$Species,
     pch = 19,
     main = "Iris Dataset",
     sub = "Sepal Width and Length",
     xlab = "Sepal Length",
     ylab = "Sepal Width")

legend("topright", legend = levels(iris$Species), col = 1:3, pch = 19)
```
