---
title: "RWorksheet_camasa#4c"
author: "April Marie"
date: "2025-01-04"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

library(ggplot2)
library(dplyr)

#1.
#1a. How to Import a CSV File
```{r}
mpgdata <- read.csv("mpg.csv")
```
#b.
```{r}
str(mpgdata)
```
2. Which manufacturer has the most models in this data set? Which model has the most variations?

```{r}
library(dplyr)
manufacturer_count <- mpgdata %>%
  group_by(manufacturer) %>%
  summarise(model_count = n_distinct(model)) %>%
  arrange(desc(model_count))
most_models_manufacturer <- manufacturer_count[1, ]
model_variation_count <- mpgdata %>%
  group_by(model) %>%
  summarise(variation_count = n()) %>%
  arrange(desc(variation_count))
most_variations_model <- model_variation_count[1, ]
most_models_manufacturer
most_variations_model
```
  A.
```{r}
unique_models <- mpgdata %>%
  group_by(manufacturer) %>%
  summarise(unique_models = list(unique(model))) %>%
  arrange(manufacturer)
print(unique_models)
```
  B. 
```{r}
model_counts <- mpgdata %>%
  group_by(manufacturer) %>%
  summarise(unique_model_count = n_distinct(model)) %>%
  arrange(desc(unique_model_count))
barplot(model_counts$unique_model_count, 
        names.arg = model_counts$manufacturer, 
        las = 2, 
        col = "lightblue", 
        main = "Number of Unique Models by Manufacturer",
        xlab = "Manufacturer", 
        ylab = "Number of Unique Models")
library(ggplot2)
ggplot(model_counts, aes(x = reorder(manufacturer, unique_model_count), y = unique_model_count)) +
  geom_bar(stat = "identity", fill = "lightblue") +
  coord_flip() +
  labs(title = "Number of Unique Models by Manufacturer", 
       x = "Manufacturer", 
       y = "Number of Unique Models") +
  theme_minimal()
```
2. Same dataset will be used. You are going to show the relationship of the modeland the manufacturer.

a. What does ggplot(mpg, aes(model, manufacturer)) + geom_point() show?

-the x-axis represents the model, and the y-axis represents the manufacturer. Since both variables are categorical, the plot will display points for each model-manufacturer pair. This can result in many overlapping points, making it difficult to identify any clear patterns or relationships. As a result, the plot may look cluttered and may not effectively convey the connection between these variables.


b. For you, is it useful? If not, how could you modify the data to make it more informative?

-As it is, the plot does not effectively reveal the relationship between the model and manufacturer due to the overplotting of points and the lack of numerical or continuous variables that could provide further insights.

3. Plot the model and the year using ggplot(). Use only the top 20 observations. Write the codes and its results.


```{r}
library(ggplot2)
library(dplyr)
top_20_mpg <- mpgdata %>%
  arrange(desc(cty)) %>%
  head(20)
ggplot(top_20_mpg, aes(x = model, y = year)) +
  geom_point() +
  labs(title = "Top 20 Models by Year", x = "Model", y = "Year") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

4. Using the pipe (%>%), group the model and get the number of cars per model. Show codes and its result.


```{r}
model_counts <- mpg %>%
  group_by(model) %>%
  summarise(number_of_cars = n()) %>%
  arrange(desc(number_of_cars))
print(model_counts)
```
  
  A.
  
```{r}
top_models <- mpgdata %>%
  group_by(model) %>%
  summarise(number_of_cars = n()) %>%
  arrange(desc(number_of_cars)) %>%
  slice_head(n = 20)
ggplot(top_models, aes(x = reorder(model, number_of_cars), y = number_of_cars, fill = model)) +
  geom_bar(stat = "identity") +
  coord_flip() +  
  labs(title = "Top 20 Car Models by Number of Cars",
       x = "Car Model",
       y = "Number of Cars",
       fill = "Model") +
  theme_minimal() +
  scale_fill_viridis_d() 
```


b. Plot using the geom_bar() + coord_flip() just like what is shown below. Show codes and its result.


```{r}
library(viridis)
library(viridisLite)
top_models <- mpgdata %>%
  group_by(model) %>%
  summarise(number_of_cars = n()) %>%
  arrange(desc(number_of_cars)) %>%
  slice_head(n = 20)
ggplot(top_models, aes(x = reorder(model, number_of_cars), y = number_of_cars, fill = model)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Top 20 Car Models by Number of Cars",
       x = "Car Model",
       y = "Number of Cars") +
  theme_minimal() +
  scale_fill_viridis_d()
```
 
 5.
 
```{r}
ggplot(mpgdata, aes(x = cyl, y = displ, color = displ)) +
  geom_point(size = 3, alpha = 0.7) +  
  labs(title = "Relationship between No. of Cylinders and Engine Displacement",
       x = "Number of Cylinders",
       y = "Engine Displacement") +
  theme_minimal() +
  scale_color_viridis_c() 
```

a. How would you describe its relationship? Show the codes and its result.
-The scatter plot shows a positive correlation between cylinder count and engine displacement, with higher cylinder counts generally linked to greater displacement. Clusters may form by cylinder count, and there can be variability, especially at higher counts. Outliers, such as high displacement with a low cylinder count, may suggest specialized engines.
6. Plot the relationship between displ (engine displacement) and hwy(highway miles per gallon). Mapped
it with a continuous variable you have identified in #1-c. What is its result? Why it produced such
output?
```{r}
ggplot(mpg, aes(x = displ, y = hwy, color = cty)) +
  geom_point(size = 3, alpha = 0.7) + 
  labs(title = "Relationship between Engine Displacement and Highway MPG",
       x = "Engine Displacement (L)",
       y = "Highway MPG",
       color = "City MPG") +
  theme_minimal() +
  scale_color_viridis_c()
```
6. Import the traffic.csv onto your R environment.

```{r}
library(readr)
traffic_data <- read_csv("traffic.csv")
head(traffic_data)
```

 A.
How many numbers of observation does it have? What are the variables of the traffic dataset.

```{r}
num_observations <- nrow(traffic_data)
variables <- colnames(traffic_data)
num_observations
variables
```
 B.
subset the traffic dataset into junctions. What is the R codes and its output?

```{r}
junctions_data <- subset(traffic_data, Junction == TRUE)
head(junctions_data)
```

 C.
Plot each junction in a using geom_line(). Show your solution and output.


```{r}
library(ggplot2)
ggplot(traffic_data, aes(x = DateTime, y = Vehicles, color = factor(Junction))) +
  geom_line() +
  labs(title = "Vehicle Counts at Junctions Over Time",
       x = "Date and Time",
       y = "Number of Vehicles",
       color = "Junction") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

  7.
 A.
How many observations does alexa_file has? What about the number of columns? Show your solution and answer.
```{r}
library(readxl)
alexa_data <- read_excel("alexa_file.xlsx")
dimensions <- dim(alexa_data)
num_rows <- dimensions[1]
num_columns <- dimensions[2]
num_rows
num_columns  
```
 B.
group the variations and get the total of each variations. Use dplyr package. Show solution and
answer.
```{r}
library(dplyr)
variation_totals <- alexa_data %>%
  group_by(variation) %>%
  summarise(total = n())
variation_totals
```

 C. 
 Plot the variations using the ggplot() function. What did you observe? Complete the details of the graph. 
 
```{r}

library(ggplot2)
ggplot(variation_totals, aes(x = variation, y = total, fill = variation)) +
  geom_bar(stat = "identity") +
  labs(title = "Total Counts for Each Variation", x = "Variation", y = "Total Count") +
  theme_minimal()
```

  D.  
Plot a geom_line() with the date and the number of verified reviews. Complete the details of the graphs.

```{r}
alexa_data$verified_reviews <- as.numeric(alexa_data$verified_reviews)
reviews_by_date <- alexa_data %>%
  group_by(date) %>%
  summarise(total_verified_reviews = sum(verified_reviews, na.rm = TRUE))
library(ggplot2)
ggplot(reviews_by_date, aes(x = date, y = total_verified_reviews)) +
  geom_line(color = "blue") +
  labs(title = "Verified Reviews Over Time", x = "Date", y = "Number of Verified Reviews") +
  theme_minimal()
```

  E.
Get the relationship of variations and ratings. Which variations got the most highest in rating? 

Plot a graph to show its relationship. Show your solution and answer.
```{r}
variation_ratings <- alexa_data %>%
  group_by(variation) %>%
  summarise(average_rating = mean(rating, na.rm = TRUE))
ggplot(variation_ratings, aes(x = variation, y = average_rating, fill = variation)) +
  geom_bar(stat = "identity") +
  labs(title = "Average Rating by Variation", x = "Variation", y = "Average Rating") +
  theme_minimal()
```
