---
title: "RWorksheet_camasa#4a"
author: "April Marie"
date: "2024-10-23"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. The table below shows the data about shoe size and height. Create a data frame.
```{r}
df <- data.frame(
  
  Shoe_Size = c(6.5, 9.0, 8.5, 8.5, 10.5, 7.0, 9.5, 9.0, 13.0, 7.5, 10.5, 8.5, 12.0,               10.5,13.0, 11.5, 8.5, 5.0, 10.0, 6.5, 7.5, 8.5, 10.5, 8.5, 10.5, 11.0, 9.0,  13.0),
  
  Height = c(66.0, 68.0, 64.5, 65.0, 70.0, 64.0, 70.0, 71.0, 72.0, 64.0, 74.5, 67.0,               71.0, 71.0, 77.0, 72.0, 59.0, 62.0, 72.0, 66.0, 64.0, 67.0, 73.0, 69.0, 72.0, 70.0, 69.0, 70.0),
  
  Gender = c("F", "F", "F", "F", "M", "F", "F", "F", "M", "F", "M", "F", "M", "M", "M", "M", "F", "F", "M", "F", "F", "M", "M", "F", "M", "M", "M", "M")
)
df
```

# a. Describe the data.
# The data contains two sets of observations for shoe size, height, and gender.

# b. Create a subset by males and females with their corresponding shoe size and height. What its result? Show the R scripts.

```{r}
males <- df[df$Gender == "M", c("Shoe_Size", "Height")]
females <- df[df$Gender == "F", c("Shoe_Size", "Height")]

males
females
```

# c. Find the mean of shoe size and height of the respondents. Write the R scripts and its result.
```{r}
mean_shoe_size <- mean(df$Shoe_Size)
mean_height <- mean(df$Height)

mean_shoe_size
mean_height
```

# d. Is there a relationship between shoe size and height? Why?

```{r}
correlation <- cor(df$Shoe_Size, df$Height)
correlation
```

# 2. Construct character vector months to a factor with factor() and assign the result to factor_months_vector. Print out factor_months_vector and assert that R prints out the factor levels below the actual values.
```{r}
months_vector <- c(
"March", "April", "January", "November", "January", "September", "October", "September", "November", "August", "January", "November", "November", "February", "May", "August", "July", "December", "August", "August", "September", "November","February", "April")
months_vector

factor_months_vector <- factor(months_vector)
factor_months_vector
```

# 3. Then check the summary() of the months_vector and factor_months_vector. | Interpret the results of both vectors. Are they both equally useful in this case?
```{r}
summary(months_vector)
summary(factor_months_vector)
```

# 4. Create a vector and factor for the table below.
```{r}
directions_vector <- c("East", "West", "North")
frequencies_vector <- c(1, 4, 3)
factor_data <- factor(directions_vector)

new_order_data <- factor(factor_data, levels = c("East", "West", "North"))
new_order_data
```

# 5. Enter the data below in Excel with file name = import_march.csv. 
# a. Import the excel file into the Environment Pane using read.table() function. Write the code.
```{r}
data <- read.table("import_march.csv", header = TRUE, sep = ",")
```

# b. View the dataset. Write the R scripts and its result.
```{r}
data
```

# 6. Full Search
```{r}
user =(readline(prompt = "Enter any number from 1 to 50: "))


if(user == 20){
  print(TRUE)
}else if(user >= 1 && user <= 50){
  print(user)
}else{
  print("The number selected is beyond the range of 1 to 50")
}
```

# 7. Change
```{r}
min_bills <- function(price) {
  bills <- c(1000, 500, 200, 100, 50)
  count <- 0
  for (bill in bills) {
    
    while (price >= bill) {
      price <- price - bill  
      count <- count + 1   
    }
  }
  return(count)
}
snack_price <- 300
cat("Minimum number of bills needed:", min_bills(snack_price), "\n")
```

# 8. The following is each student’s math score for one semester. Based on this, answer the following questions.
# a. Create a dataframe from the above table. Write the R codes and its output.
```{r}
students_data <- data.frame(
  Name = c("Annie", "Thea", "Steve", "Hanna"),
  Grade1 = c(85, 65, 75, 95),
  Grade2 = c(65, 75, 55, 75),
  Grade3 = c(85, 90, 80, 100),
  Grade4 = c(100, 90, 85, 90)
)

print(students_data)
```

# b. Without using the rowMean function, output the average score of students whose average math score over 90 points during the semester. write R code and its output.
```{r}
calculate_average <- function(grades) {
  total <- sum(grades)
  average <- total / length(grades)
  return(average)
}
for (i in 1:nrow(students_data)) {
  grades <- as.numeric(students_data[i, 2:5])
  average_grade <- calculate_average(grades)
  
  if (average_grade > 90) {
    cat(students_data$Name[i], "'s average grade this semester is ", average_grade, "\n", sep = "")
  }
}
```
# c. Without using the mean function, output as follows for the tests in which the average score was less than 80 out of 4 tests.
```{r}
for (j in 2:ncol(students_data)) {
  total <- sum(students_data[, j])
  avg_test <- total / nrow(students_data)
  
  if (avg_test < 80) {
    cat("The ", j - 1, "nd test was difficult.\n", sep = "")
  }
}
```
# d. Without using the max function, output as follows for students whose highest score for a semester exceeds 90 points.
```{r}
calculate_max <- function(grades) {
  max_grade <- grades[1]
  for (grade in grades) {
    if (grade > max_grade) {
      max_grade <- grade
    }
  }
  return(max_grade)
}

for (i in 1:nrow(students_data)) {
  grades <- as.numeric(students_data[i, 2:5])
  highest_grade <- calculate_max(grades)
  
  if (highest_grade > 90) {
    cat(students_data$Name[i], "'s highest grade this semester is ", highest_grade, "\n", sep = "")
  }
}
```