---
title: "RWorksheet_camasa#3"
author: "April Marie"
date: "2024-10-02"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Using VECTORS
# 1.
# a. You need to produce a vector that contains the first 11 letters.
```{r}
first11_letters <- LETTERS[1:11]
first11_letters
```
# b. Produce a vector that contains the odd numbered letters.
```{r}
oddNumber_letters <- LETTERS[seq(1, 26, by=2)]
oddNumber_letters
```
# c. Produce a vector that contains the vowels.
```{r}
vowels <- LETTERS[c(1, 5, 9, 15, 21)] 
vowels
```
# d. Produce a vector that contains the last 5 lowercase letters.
```{r}
last5_lowercase <- letters[22:26]
last5_lowercase
```
# e. Produce a vector that contains letters between 15 to 24 letters in lowercase.
```{r}
lowercase15_to_24 <- letters[15:24]
lowercase15_to_24
```

# 2.
# a. Name the object as city.
```{r}
city <- c("Tuguegarao City", "Manila", "Iloilo City", "Tacloban", "Samal Island", "Davao City")
city
```
# b. Name the object as temp. Write the R code and its output.
```{r}
temp <- c(42, 39, 34, 34, 30, 27)
temp
```
# c. Create a dataframe to combine the city and the temp by using ‘data.frame(). What the R code and its result?
```{r}
df_cityTemp <- data.frame(city, temp)
df_cityTemp
```
# d. Change the column names by using names() function as City and Temperature. What is the R code and its result?
```{r}
names(df_cityTemp) <- c("City", "Temperature")
df_cityTemp
```
# e. Print the structure by using str() function. Describe the output.
```{r}
str(df_cityTemp)
```
# f. From the answer in d, what is the content of row 3 and row 4 What is its R code and its output?
```{r}
df_cityTemp[3:4,]
```
# g.
```{r}
highestTempCity <- df_cityTemp$City[which.max(df_cityTemp$Temperature)]
lowestTempCity <- df_cityTemp$City[which.min(df_cityTemp$Temperature)]

highestTempCity
lowestTempCity
```

# 2. using MATRICES
# a. What will be the R code for the #2 question and its result?
```{r}
matr <- matrix(c(1:8, 11:14), nrow = 3, ncol = 4)
matr
```
# b. Multiply the matrix by two. What is its R code and its result?
```{r}
multiply <- matr * 2
multiply
```
# c. What is the content of row 2? What is its R code?
```{r}
matr[2, ]
```
# d. What will be the R code if you want to display the column 3 and column 4 in row 1 and row 2? What is its output?
```{r}
matr[1:2, 3:4]
```
# e. What is the R code is you want to display only the columns in 2 and 3, row 3? What is its output?
```{r}
matr[3, 2:3]
```
# f. What is the R code is you want to display only the columns 4? What is its output?
```{r}
matr[ ,4]
```
# g. Name the rows as isa, dalawa, tatlo and columns as uno, dos, tres, quatro for the matrix that was created in b.‘. What is its R code and corresponding output?
```{r}
rownames(matr) <- c("isa", "dalawa", "tatlo")
colnames(matr) <- c("uno", "dos", "tres", "quatro")
matr
```
# h. New dimensions should have 2 columns and 6 rows. What will be the R code and its output?
```{r}
dim(matr) <- c(6, 2)
matr
```
# 3. using ARRAYS
# a. numeric values.
```{r}
array1 <- rep(c(1, 2, 3, 6, 7, 8, 9, 0, 3, 4, 5, 1), 2)
numarray <- array(array1, dim = c(2, 4, 3))
numarray
# b. How many dimensions do your array have?
dim(numarray)
# c.
rownames(numarray) <- c("a", "b")
colnames(numarray) <- c("A", "B", "C", "D")
dimnames(numarray)[[3]] <- c("1st-Dimensional Array", "2nd-Dimensional Array", "3rd-Dimensional Array")

numarray
```
