---
title: "RWorksheet_camasa#2"
author: "April Marie"
date: "2024-09-25"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1.Create a vector using : operator
# a. Sequence from -5 to 5. Write the R code and its output.
```{r}
operator <- c(-5:5)
print(operator)
```
# The output is sequenced from -5 to 5.

# b. x <- 1:7. What will be the value of x?
```{r}
x <- seq(1:7)
print(x)
```
# The values of x are 1 2 3 4 5 6 7.

# 2. Create a vector using seq() function
# a. seq(1, 3, by=0.2) # specify step size
```{r}
seqC <- seq(1, 3, by=0.2)
print(seqC)
```
# The output is a numeric vector that starts at 1 and ends at 3, with increments (or step size) of 0.2.

# 3. A factory has a census of its workers. There are 50 workers in total. The following
list shows their ages: 34, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27, 22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 43, 53, 41, 51, 35, 24,33, 41, 53, 40, 18, 44, 38, 41, 48, 27, 39, 19, 30, 61, 54, 58, 26, 18.
# a. Access 3rd element, what is the value?
```{r}
ages <- c(34, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27, 22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 43, 53, 41, 51, 35, 24, 33, 41, 53, 40, 18, 44, 38, 41, 48, 27, 39, 19, 30, 61, 54, 58, 26, 18)
third <- ages[3]
print(third)
```
# b. Access 2nd and 4th element, what are the values?
```{r}
second_fourth <- ages[c(2, 4)]
print(second_fourth)
```
# c. Access all but the 4th and 12th element is not included. Write the R script and its output.
```{r}  
age <- ages[-c(4, 12)]
print(age)
```
# This output is the original list of ages without the 4th and 12th elements.

# 4. Create a vector x <- c("first"=3, "second"=0, "third"=9). Then named the vector, names(x).
```{r}
names <- c("first"=3, "second"=0, "third"=9)
print(names)
```
# a. Print the results. Then access x[c("first", "third")]. Describe the output.
```{r}
names[c("first", "third")]
```
# The output prints the first and third.

# 5. Create a sequence x from -3:2.
```{r}
seqA <- seq(-3:2)
print(seqA)
```
# a. Modify 2nd element and change it to 0;
#x[2] <- 0 
#x. 
#Describe the output.
```{r}
x[2] <- 0 
x
```

# 6. The following data shows the diesel fuel purchased by Mr. Cruz.
# a. Create a data frame for month, price per liter (php) and purchase-quantity (liter). Write the R scripts and its output.
```{r}
months <- c("January", "February", "March", "April", "May", "June")
price_per_liter <- c(52.50, 57.25, 60.00, 65.00, 74.25, 54.00)
purchase_quantity <- c(25, 30, 40, 50, 10, 45)

data <- data.frame(Month = months, Price_Per_Liter = price_per_liter, Purchase_Quantity = purchase_quantity)
print(data)
```
# b. What is the average fuel expenditure of Mr. Cruz from Jan to June? Note: Use ‘weighted.mean(liter, purchase)‘. Write the R scripts and its output.
```{r}
expenditure <- price_per_liter * purchase_quantity

average_expenditure <- weighted.mean(expenditure, purchase_quantity)

print(average_expenditure)
```

# 7. R has actually lots of built-in datasets. For example, the rivers data “gives the lengths (in miles) of 141 “major” rivers in North America, as compiled by the US Geological Survey”.
# Type “rivers” in your R console.
```{r}
data <- c(length(rivers), sum(rivers), mean(rivers), median(rivers), var(rivers), sd(rivers), min(rivers), max(rivers))
print(data)
```

